---
title: "STAT40830 - Homework 1"
author: "Shreemadhi Babu Rajendra Prasad"
format: pdf
execute:
  echo: false
editor: visual
---

## Introduction

The **'diamonds'** is an in built dataset from **\`ggplot2\`** package which contains *pricing and quality* information of around 50,000 diamonds.

Each row represents the data of a single diamond with the key variables:

-   **carat**: *Weight of the diamond*
-   **cut**: *Quality of the cut (Fair, Good, Very Good, Premium, Ideal)*
-   **color**: *Diamond color, from D (best) to J (worst)*
-   **clarity**: *Measurement of internal flaws*
-   **depth**: *Total depth percentage = z / mean(x, y)*
-   **table**: *Width of the top of the diamond relative to its widest point*
-   **price**: *Price in US dollars*
-   **x, y, z**: *Length, width, and depth (in mm)*

This dataset is commonly used for demonstrate data visualization, statistical modeling, etc.

```{r}
library(ggplot2)
dim(diamonds)
```

The dataset consists of **53940 rows** and **10 columns**.

## **Diamond Price Distribution by Clarity** [(Boxplot)]{style="color:darkblue"}

```{r}
ggplot(diamonds, aes(x = clarity, y = price)) +
  geom_boxplot(fill = "darkred") +
  labs(
    title = "Diamond Price Distribution by Clarity",
    x = "Clarity",
    y = "Price (USD)"
  ) +
  theme_minimal()
```

This **boxplot** shows a variation between *diamond prices* across different **clarity levels**.

-   The median price differs across the **clarity types**.
-   Clarity grades like IF and VVS1 shows a **lower median prices**, despite being higher quality.
-   Diamonds with lower clarity (e.g., I1, SI2) often show **higher median prices** due to larger carat sizes.
-   There are **many outliers** — especially in lower clarity grades which indicates that **price is influenced by multiple factors**, not just clarity.
